---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# dsdataprep


<!-- badges: start -->
<!-- badges: end -->

<br>


## Overview

dsdataprep helps assisting data visualization packages by offering functions to aggregate, summarise, sort and wrap data; adds color to the data columns and adds rows information for displaying in interactive plots

### Main functions:

- `data_aggregation()`

- `wrap_sort_data()`

- `prep_tooltip()`

- `add_data_colors()`



## Installation

Install the development version of makeup from GitHub with:

``` r
# install.packages("devtools")
remotes::install_github("datasketch/dsdataprep")
```


## Example

This is a basic example which shows you how this packages work:

Let´s load `dsdataprep` package

```{r library}
library(dsdataprep)
```


### Aggregate and summarie data with `aggregation_data()` function

`aggregation_data()` allows you to summaries information (e.g. count, sum, mean) and gives you options to do it by a (or more) grouping variable(s), adding custom names for the new columns, among other options.


#### Counting values
```{r}
# Load data
data <- ggplot2::diamonds

# Aggregate
data_result <- aggregation_data(data = data,
                                 agg = "count",
                                 to_agg = NULL,
                                 agg_name = "Conteo",
                                 group_var = c("cut", "color"))

data_result
```


#### Sum values by a grouping variable and add custom names for new columns
```{r}
data_result <- aggregation_data(data = data,
                                agg = "sum",
                                to_agg = c("x", "y"),
                                agg_name = c("Sum x", "Sum y"),
                                group_var = c("cut"),
                                percentage = TRUE)

data_result
```


### Sort and wrap data by a categorical variable and numeric variable with `wrap_sort_data()`:

```{r}
# Load data
data <- ggplot2::diamonds

# Checking
data[!duplicated(data$cut), "cut"]


# Sort data
data_result <- wrap_sort_data(data, 
                              col_cat = "cut", 
                              order = c("Good", "Ideal"))

# Checking
data_result[!duplicated(data_result$cut), "cut"]
```


### Generate a tooltip string for each row of a data frame or matrix with `prep_tooltip()`. Tooltips are typically used to display additional information about the data when the user hovers over a data point in a plot or table:

```{r}
# Load data
data <- ggplot2::diamonds

# Add tooltip
v <- prep_tooltip(data,
                    format_num = "1345.78",
                    opts_format_num = list(si_prefix = TRUE),
                    format_cat = "UPPER",
                    tooltip = "Precio: {price} <br/> Tipo: {cut}")
```



Learn about the many ways to work with formatting dates values in `vignette("intro-to-dsdataprep")`
